### 2025-04-30 MSSC5750_Project.qmd
Added Shap Values & more Viz code

---
title: "Final Project"
author: "David, Michael, Patrick"
format: html
---

```{r echo = FALSE, message=FALSE}
library(MASS)
library(ggplot2)
library(dplyr)
library(scales)
library(glmnet)
library(fastDummies)
library(caret)
library(shapr)
library(shapviz)
library(party)
library(reshape2)

# Importing data set from github repository
data <- read.csv("https://raw.githubusercontent.com/Pburns525/MSSC5750-Project/refs/heads/main/loan_data.csv")

```

High-Level Data Exploration
```{r}

nrow(data)
colnames(data)

# There is no missing data, already a cleaned data set
colSums(is.na(data))

table(data$person_gender)

# Education is not currently ordered, change to a factor and set levels
data$person_education<- factor(data$person_education, levels=c('High School','Associate','Bachelor','Master','Doctorate'), ordered=TRUE)
table(data$person_education)

# Looks like we have somebody who is 144 yrs old, 7 who are over age 100. PRobably want to remove them, its doubtful anybody very old would be applying anyways, and these ages are not really possible
summary(data$person_age)
sum(data$person_age>100)

summary(data$person_income)

summary(data$loan_amnt)

table(data$loan_intent)

summary(data$cb_person_cred_hist_length)

summary(data$loan_percent_income)

summary(data$credit_score)

table(data$person_home_ownership)

# This is years of employment experience. Again we see some super high values, longer than human lifespan, so should clean those up
summary(data$person_emp_exp)
sum(data$person_emp_exp>70)

table(data$previous_loan_defaults_on_file)

# 22% of our data receives a loan, so our dataset is imbalanced, but not too extreme
percent(table(data$loan_status)[2]/nrow(data),accuracy=1)

# Oddly higher credit score does not seem to make a difference
cor(data$credit_score,data$loan_status)
summary(data$credit_score)
lowcred<-data %>% filter(credit_score<601)
goodcred<-data %>% filter(credit_score>670)
mean(lowcred$loan_status)
mean(goodcred$loan_status)

```

Creating some basic visualizations
```{R}

# Removing outliers (Somebody was earning like 7M, we have almost 25 who are over 1M still)
ggplot(data=data %>% filter(person_income<500000), aes(x=person_income)) +geom_histogram(bins=30)

#Actually might be better to just add a catch bucket at the end. IF over 250k, just set to 250k
ggplot(data=data %>% mutate(person_income2 = if_else(person_income<250000,person_income,250000)), aes(x=person_income2)) +geom_histogram(bins=40, fill='steelblue',color='black') + xlab("Income") + ggtitle("Histogram of Income")+
  scale_x_continuous(breaks=seq(0,250000,50000), labels=c("$0","$50,000","$100,000","$150,000","$200,000","$250,000+") )

temp<-data %>% summarize('0'=1-mean(loan_status),"1"=mean(loan_status) ,.by=loan_intent)
intent<-melt(temp) %>% dplyr::rename(loan_status=variable)

ggplot(data=intent, aes(x=loan_intent, y=value, group=loan_status, fill=loan_status)) + geom_col(position='stack')

```

 Outline for LASSO regression to help with variable selection
```{R}

# Preparing categorical variables. Should happen automatically for factors
data$person_gender <- as.factor(data$person_gender)
data$person_home_ownership<-as.factor(data$person_home_ownership)
data$loan_intent<-as.factor(data$loan_intent)
data$previous_loan_defaults_on_file<-as.factor(data$previous_loan_defaults_on_file)

# Extracting outcome variable
class <- data$loan_status

# Converting to a matrix
matrixData <- model.matrix(class ~ ., data = data %>% select(-loan_status))

# First column is all intercepts
matrixData<- matrixData[,-1]

# Uses cross validation, no need to train/test split. Alpha selects lasso (rather than ridge regression), but I think thats the default anyways
lasso <- cv.glmnet(matrixData, class, alpha = 1, family = "binomial")

# See coefficients, which variables we might drop
coef(lasso)

# Make predictions
data$lassoPred <- as.vector(predict(lasso, newx = matrixData, type = "class"))

# Evaluate accuracy measures
confusionMatrix(factor(data$lassoPred), factor(data$loan_status), positive = '1')

```

Now doing a basic logistic regression with/without lasso selection
```{R}

# We need a train/test split. We will use 80/20
# can use uniform distribution, set seed for reproducability
set.seed(100)
forTrain<-runif(nrow(data))<.8

trainData<-data[forTrain,]
testData<-data[!forTrain,]

# First creating logistic regression with all our variables
logModel <- glm(loan_status ~ person_age + person_gender + person_education + person_income + person_emp_exp + person_home_ownership +loan_amnt + loan_intent + loan_int_rate + loan_percent_income + cb_person_cred_hist_length + credit_score + previous_loan_defaults_on_file, data=trainData, family= binomial )

# Repeating but removing the variables not relevant in lasso
# removed(person cred length, person_emp_exp person_income person_gender person_education person_age)
logModelLasso <- glm(loan_status ~ person_home_ownership +loan_amnt + loan_intent + loan_int_rate + loan_percent_income  + credit_score + previous_loan_defaults_on_file, data=trainData, family= binomial )

testData$logisticPred <- round(predict(logModel, testData, type='response'))
testData$logisticLassoPred <- round(predict(logModelLasso, testData, type='response'))

# Matrix for regular log
confusionMatrix(factor(testData$logisticPred), factor(testData$loan_status), positive = '1')

# Matrix for lasso log
confusionMatrix(factor(testData$logisticLassoPred), factor(testData$loan_status), positive = '1')

```

Shap values/viz
```{r}
# Requires split out xtrain, ytrain, xtest... Also a combined x/ytrain

# We will apply to lasso data

# We need to manually generate the dummy variables and match naems to what the model created
# coder<-dummyVars(~ person_home_ownership + loan_intent + previous_loan_defaults_on_file, data = data, fullRank = TRUE, sep='')
# trainDataAdj<-cbind(trainData %>% select(-person_home_ownership, -loan_intent, -previous_loan_defaults_on_file), predict(coder, newdata = trainData) )
# testDataAdj<-cbind(testData %>% select(-person_home_ownership, -loan_intent, -previous_loan_defaults_on_file), predict(coder, newdata = testData) )

# Only want the columns that we included in lasso

xtrain<-trainData %>% select(person_home_ownership ,loan_amnt , loan_intent , loan_int_rate , loan_percent_income  , credit_score , previous_loan_defaults_on_file)
ytrain<-trainData %>% select(loan_status)
xytrain<-trainData %>% select(person_home_ownership ,loan_amnt , loan_intent , loan_int_rate , loan_percent_income  , credit_score , previous_loan_defaults_on_file,loan_status)
xtest<-testData  %>% select(person_home_ownership ,loan_amnt , loan_intent , loan_int_rate , loan_percent_income  , credit_score , previous_loan_defaults_on_file)

# calculating shap is costly so we will reduce the size of the data we pass in

# we will just explain on our train data as well for larger sample size
set.seed(5)
xtrainSmall <- xtrain[sample(nrow(xtrain), 1000), ]
explainer<- shapr::explain(model=logModelLasso, x_train=xtrainSmall, x_explain=xtrainSmall, approach="ctree", phi0=mean(ytrain$loan_status))

shp <- shapviz(explainer)
sv_importance(shp)

plot(explainer, plot_phi0 = FALSE, index_x_explain = 1)

shapObj <-shapviz(explainer)
sv_importance(shapObj, kind="beeswarm", show_numbers = TRUE, max_display=15)

```