### 2025-05-01 MSSC5750_Project.qmd
Data Cleaning, One-Hot-Encoding, Results Table

---
title: "Final Project"
author: "David, Michael, Patrick"
format: html
---

```{r echo = FALSE, message=FALSE}
library(MASS)
library(ggplot2)
library(dplyr)
library(scales)
library(glmnet)
library(fastDummies)
library(caret)
library(shapr)
library(shapviz)
library(party)
library(reshape2)

# Importing data set from github repository
data <- read.csv("https://raw.githubusercontent.com/Pburns525/MSSC5750-Project/refs/heads/main/loan_data.csv")

```

High-Level Data Exploration
```{r}

nrow(data)
colnames(data)

# There is no missing data, already a cleaned data set
colSums(is.na(data))

table(data$person_gender)

# Education is not currently ordered, change to a factor and set levels
data$person_education<- factor(data$person_education, levels=c('High School','Associate','Bachelor','Master','Doctorate'), ordered=TRUE)
table(data$person_education)

# Looks like we have somebody who is 144 yrs old, 7 who are over age 100. PRobably want to remove them, its doubtful anybody very old would be applying anyways, and these ages are not really possible
summary(data$person_age)
sum(data$person_age>100)

summary(data$person_income)

summary(data$loan_amnt)

table(data$loan_intent)

summary(data$cb_person_cred_hist_length)

summary(data$loan_percent_income)

summary(data$credit_score)

table(data$person_home_ownership)

# This is years of employment experience. Again we see some super high values, longer than human lifespan, so should clean those up
summary(data$person_emp_exp)
sum(data$person_emp_exp>70)

# Hm. In every single case where there was a prior loan default, loan_status=0. We probably don't want to include this variable in anything explanatory
table(data$previous_loan_defaults_on_file, data$loan_status)

# 22% of our data receives a loan, so our dataset is imbalanced, but not too extreme
percent(table(data$loan_status)[2]/nrow(data),accuracy=1)

# Oddly higher credit score does not seem to make a difference
cor(data$credit_score,data$loan_status)
summary(data$credit_score)
lowcred<-data %>% filter(credit_score<601)
goodcred<-data %>% filter(credit_score>670)
mean(lowcred$loan_status)
mean(goodcred$loan_status)

```

Creating some basic visualizations
```{R}

# Removing outliers (Somebody was earning like 7M, we have almost 25 who are over 1M still)
ggplot(data=data %>% filter(person_income<500000), aes(x=person_income)) +geom_histogram(bins=30)

#Actually might be better to just add a catch bucket at the end. IF over 250k, just set to 250k
ggplot(data=data %>% mutate(person_income2 = if_else(person_income<250000,person_income,250000)), aes(x=person_income2)) +geom_histogram(bins=40, fill='steelblue',color='black') + xlab("Income") + ggtitle("Histogram of Income")+
  scale_x_continuous(breaks=seq(0,250000,50000), labels=c("$0","$50,000","$100,000","$150,000","$200,000","$250,000+") )

temp<-data %>% summarize('0'=1-mean(loan_status),"1"=mean(loan_status) ,.by=loan_intent)
intent<-melt(temp) %>% dplyr::rename(loan_status=variable)

ggplot(data=intent, aes(x=loan_intent, y=value, group=loan_status, fill=loan_status)) + geom_col(position='stack')

```

Data cleaning
```{R}

data2<-data
######## Transformation of variables ########

#### person_income ####

# Salary is extremely right skewed, lets use the log transformation. Minimum is 8000 so we do not need to add a constant
min(data2$person_income)
data2$log_person_income <- log(data2$person_income)

# Our max is still way larger than our 3rd quartile, so maybe we should cap salary before transformation
summary(data2$log_person_income)
data2$log_person_income<-log(ifelse(data2$person_income>500000,500000,data2$person_income))

# Better
summary(data2$log_person_income)

#### person_emp_exp ####

# Employee experience also very skewed. And some of the values dont really make any sense. Log transformation can fix this too
summary(data2$person_emp_exp)

# First replace extreme (impossible) values with something realistic
data2$person_emp_exp<-ifelse(data2$person_emp_exp>50,50,data2$person_emp_exp)

data2$log_person_emp_exp <- log(data2$person_emp_exp+.5)

# Better
summary(data2$log_person_emp_exp)

#### Age  ####
# We need to bucket our age. Theres no way we actually have someone aged 144, and 75% of our data falls between 20-30
data2<- data2 %>% mutate(person_age = case_when(person_age<25 ~ "<25",
                                                person_age<=30 ~"25-30",
                                                person_age<45 ~ "31-44",
                                                person_age<60 ~ "45-59",
                                                TRUE ~ "60+"))

data2$person_age<-as.factor(data2$person_age)
table(data2$person_age)

####  cred hist length  ####
# Only a handful fall >=18, we will just cap it all there as a "catch all", but leave as numeric
table(data2$cb_person_cred_hist_length)
data2$cb_person_cred_hist_length <- ifelse(data2$cb_person_cred_hist_length>18,18, data2$cb_person_cred_hist_length)

######## One hot encoding for categorical variables ########

# Preparing categorical variables.
data2$person_gender <- as.factor(data2$person_gender)
data2$person_home_ownership<-as.factor(data2$person_home_ownership)
data2$loan_intent<-as.factor(data2$loan_intent)
data2$previous_loan_defaults_on_file<-as.factor(data2$previous_loan_defaults_on_file)

# One hot encoding for categorical variables
data3 <- dummy_cols(data2, remove_first_dummy = FALSE, remove_selected_columns = TRUE)

# we should remove first dummy to prevent multicolinnearity , need to keep track of what was dropped because that is our reference value
data3<-data3 %>% select(-c("person_gender_female","person_education_High School","person_home_ownership_MORTGAGE", "loan_intent_DEBTCONSOLIDATION", "previous_loan_defaults_on_file_No", "person_age_<25"))

# replacing base data with cleaner version now
cleanData <- data3

```

 Outline for LASSO regression to help with variable selection
```{R}

# Extracting outcome variable
class <- cleanData$loan_status

# Converting to a matrix, exclude the non-log-transformed, and also the prior loan default
matrixData <- model.matrix(class ~ ., data = cleanData %>% select(-c("loan_status", "person_income", "person_emp_exp", "previous_loan_defaults_on_file_Yes") ))

# First column is all intercepts
matrixData<- matrixData[,-1]

# Uses cross validation, no need to train/test split. Alpha selects lasso (rather than ridge regression), but I think thats the default anyways
lasso <- cv.glmnet(matrixData, class, alpha = 1, family = "binomial")

# See coefficients, which variables we might drop
# Now that we switched to log income, variable is no longer being dropped. We will proceed with log_income
coef(lasso)

# Make predictions
cleanData$lassoPred <- as.vector(predict(lasso, newx = matrixData, type = "class"))

# Evaluate accuracy measures
confusionMatrix(factor(cleanData$lassoPred), factor(cleanData$loan_status), positive = '1')

# And again with our loan_default knowledge:
confusionMatrix(factor( ifelse(cleanData$previous_loan_defaults_on_file_Yes==1,0,cleanData$lassoPred)), factor(cleanData$loan_status), positive = '1')

```

Now doing a basic logistic regression with/without lasso selection
```{R}

# We need a train/test split. We will use 80/20
# can use uniform distribution, set seed for reproducability
set.seed(100)
forTrain<-runif(nrow(cleanData))<.8

trainData<-cleanData[forTrain,]
testData<-cleanData[!forTrain,]

# First creating logistic regression with all our variables
logModel <- glm(loan_status ~ . , data=trainData %>% select(-c("lassoPred", "person_income", "person_emp_exp", "previous_loan_defaults_on_file_Yes")), family= binomial )

# Repeating but only with variables selected by lasso
logModelLasso <- glm(loan_status ~ loan_int_rate + loan_percent_income + log_person_income + person_home_ownership_OTHER + person_home_ownership_OWN + person_home_ownership_RENT + loan_intent_EDUCATION + loan_intent_HOMEIMPROVEMENT + loan_intent_PERSONAL + loan_intent_VENTURE, data=trainData, family= binomial )

testData$logisticPred <- round(predict(logModel, testData, type='response'))
testData$logisticLassoPred <- round(predict(logModelLasso, testData, type='response'))

# Matrix for regular log
confusionMatrix(factor(testData$logisticPred), factor(testData$loan_status), positive = '1')

# Matrix for lasso log
confusionMatrix(factor(testData$logisticLassoPred), factor(testData$loan_status), positive = '1')

# Our training data is still heavily imbalanced, repeating again for rebalanced data
table(trainData$loan_status)
ones<-trainData %>% filter(loan_status==1)
set.seed(5)
zeros<-trainData %>% filter(loan_status==0) %>% sample_n(8000)
trainDataBalanced<-rbind(ones,zeros)
table(trainDataBalanced$loan_status)

rebalancedLog <- glm(loan_status ~ . , data=trainDataBalanced %>% select(-c("lassoPred", "person_income", "person_emp_exp", "previous_loan_defaults_on_file_Yes")), family= binomial )

testData$rebalancedLog <- round(predict(rebalancedLog, testData, type='response'))

# Matrix for regular log
confusionMatrix(factor(testData$rebalancedLog), factor(testData$loan_status), positive = '1')

```

Creating dataframe of results
```{R}

results<-data.frame(model=character(),accuracy=numeric(),sensitivity=numeric(),specificity=numeric())

addresults<-function(name,res){
  results<<-rbind(results, data.frame(model=name,accuracy=res$overall['Accuracy'],sensitivity=res$byClass['Sensitivity'],specificity=res$byClass['Specificity']) )
}

addresults('Lasso', confusionMatrix(factor( cleanData$lassoPred), factor(cleanData$loan_status), positive = '1'))
addresults('Log', confusionMatrix(factor(testData$logisticPred), factor(testData$loan_status), positive = '1'))
addresults('LogLasso', confusionMatrix(factor(testData$logisticLassoPred), factor(testData$loan_status), positive = '1'))
addresults('LogBalanced', confusionMatrix(factor(testData$rebalancedLog), factor(testData$loan_status), positive = '1'))
rownames(results)<-NULL
results

###################### Repeating for accuracy after applying logic for previous default ##############################

results2<-data.frame(model=character(),accuracy=numeric(),sensitivity=numeric(),specificity=numeric())

addresults2<-function(name,res){
  results2<<-rbind(results2, data.frame(model=name,accuracy=res$overall['Accuracy'],sensitivity=res$byClass['Sensitivity'],specificity=res$byClass['Specificity']) )
}

addresults2('Lasso', confusionMatrix(factor(ifelse(cleanData$previous_loan_defaults_on_file_Yes==1,0,cleanData$lassoPred) ), factor(cleanData$loan_status), positive = '1'))
addresults2('Log', confusionMatrix(factor( ifelse(testData$previous_loan_defaults_on_file_Yes==1,0, testData$logisticPred) ), factor(testData$loan_status), positive = '1'))
addresults2('LogLasso', confusionMatrix(factor(ifelse(testData$previous_loan_defaults_on_file_Yes==1,0,testData$logisticLassoPred) ), factor(testData$loan_status), positive = '1'))
addresults2('LogBalanced', confusionMatrix(factor(ifelse(testData$previous_loan_defaults_on_file_Yes==1,0,testData$rebalancedLog) ), factor(testData$loan_status), positive = '1'))
rownames(results2)<-NULL
results2

```

Shap values/viz
```{r}
# Requires split out xtrain, ytrain, xtest... Also a combined x/ytrain

# We will apply to lasso data

# We need to manually generate the dummy variables and match naems to what the model created
# coder<-dummyVars(~ person_home_ownership + loan_intent + previous_loan_defaults_on_file, data = data, fullRank = TRUE, sep='')
# trainDataAdj<-cbind(trainData %>% select(-person_home_ownership, -loan_intent, -previous_loan_defaults_on_file), predict(coder, newdata = trainData) )
# testDataAdj<-cbind(testData %>% select(-person_home_ownership, -loan_intent, -previous_loan_defaults_on_file), predict(coder, newdata = testData) )

# Only want the columns that we included in lasso

xtrain<-trainData %>% select(person_home_ownership ,loan_amnt , loan_intent , loan_int_rate , loan_percent_income  , credit_score , previous_loan_defaults_on_file)
ytrain<-trainData %>% select(loan_status)
xytrain<-trainData %>% select(person_home_ownership ,loan_amnt , loan_intent , loan_int_rate , loan_percent_income  , credit_score , previous_loan_defaults_on_file,loan_status)
xtest<-testData  %>% select(person_home_ownership ,loan_amnt , loan_intent , loan_int_rate , loan_percent_income  , credit_score , previous_loan_defaults_on_file)

# calculating shap is costly so we will reduce the size of the data we pass in

# we will just explain on our train data as well for larger sample size
set.seed(5)
xtrainSmall <- xtrain[sample(nrow(xtrain), 1000), ]
explainer<- shapr::explain(model=logModelLasso, x_train=xtrainSmall, x_explain=xtrainSmall, approach="ctree", phi0=mean(ytrain$loan_status))

shp <- shapviz(explainer)
sv_importance(shp)

plot(explainer, plot_phi0 = FALSE, index_x_explain = 1)
####

###
shapObj <-shapviz(explainer)
sv_importance(shapObj, kind="beeswarm", show_numbers = TRUE, max_display=15)

```